\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-22 papers/jss/article.Rnw
Add reference to jsonlite
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@815 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Update the other approaches section of conclusion with some suggestions by Phillip Yelland.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@816 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}
\shortcites{janus}
%% almost as usual
\author{Dirk Eddelbuettel\\Debian Project \And 
        Murray Stokely\\Google, Inc \And
        Jeroen Ooms\\UCLA}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

@@

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several R packages
implement functions to parse and generate \texttt{JSON} data from R
objects \citep{rjson,RJSONIO}.

@@

\section{Conclusion and Commentary}
\label{sec:summary}
% TODO(mstokely): Get cibona approval for these two sentences before
% publishing.
Schema-less text formats such as CSV and JSON will continue to be
widely used in many contexts, but we hope that the availability of
\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
pipelines to embrace schemas such as Protocol Buffers for type-safe
and performant data serialization between applications.

\pkg{RProtoBuf} has been heavily used inside Google for the past three
years by statisticians and software engineers.  At the time of this
writing there are more than XXX 30-day active users of RProtoBuf using
it to read data from and otherwise interact with other distributed
systems written in C++, Java, Python, and other languages.

\paragraph*{Other Approaches}

\pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
but it is not designed for efficient high-speed manipulation of large
numbers of protocol buffers once they have been read into R.  For
example, taking a list of 100,000 Protocol Buffers, extracting a named
field from each one, and computing an aggregate statistic on those
values would be relatively slow with RProtoBuf.  Instead for such a
use case, the current design of RProtoBuf relies on other database
systems to provide query and aggregation semantics before the
resulting protocol buffers are read into R.  Such queries could be
supported in a future version of \pkg{RProtoBuf} by supporting a
vector of messages type such that \emph{slicing} operations over a
given field across a large number of messages could be done
efficiently in C++.

>>>>

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}
\shortcites{janus,dremel}
%% almost as usual
\author{Dirk Eddelbuettel\\Debian Project \And 
        Murray Stokely\\Google, Inc \And
        Jeroen Ooms\\UCLA}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

@@

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several R packages
implement functions to parse and generate \texttt{JSON} data from R
objects \citep{rjson,RJSONIO,jsonlite}.

@@

\section{Concluding remarks}
\label{sec:summary}
% TODO(mstokely): Get cibona approval for these two sentences before
% publishing.
Schema-less text formats such as CSV and JSON will continue to be
widely used in many contexts, but we hope that the availability of
\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
pipelines to embrace schemas such as Protocol Buffers for type-safe
and performant data serialization between applications.

\pkg{RProtoBuf} has been heavily used inside Google for the past three
years by statisticians and software engineers.  At the time of this
writing there are more than XXX 30-day active users of RProtoBuf using
it to read data from and otherwise interact with other distributed
systems written in C++, Java, Python, and other languages.
\\

\emph{Other Approaches}
\\

\pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
but it is not designed for efficient high-speed manipulation of large
numbers of protocol buffers once they have been read into R.  For
example, taking a list of 100,000 Protocol Buffers, extracting a named
field from each one, and computing an aggregate statistic on those
values would be relatively slow with RProtoBuf.  Mechanisms to address
such use cases are under investigation for possible incorporation into
future releases of RProtoBuf, but currently, the package relies on
other database systems to provide query and aggregation semantics
before the resulting protocol buffers are read into R.  Inside Google,
the Dremel query system \citep{dremel} is often employed in this role
in conjunction with \pkg{RProtoBuf}.

% Such queries could be
%supported in a future version of \pkg{RProtoBuf} by supporting a
%vector of messages type such that \emph{slicing} operations over a
%given field across a large number of messages could be done
%efficiently in C++.
